Open archaeology, open source?

### 2023-10-12 analysis/_03-data_methodology.qmd
Add tbl-glossary

<!-- TODO:
* Introduce git, GitHub, etc. and describe briefly how they work for non-users
* Define software jargon we use (repo, commit, etc.) – maybe as a glossary table?
-->

>>>>

```{r tbl-glossary}
#| tbl-cap: Glossary of git and GitHub terminology
tribble(
  ~term, ~definition,
  "Comment", "Text post attached to an issue, including the first one that describes the issue",
  "Commit", "Set of changes (addition, alteration, or deletion) to files in a repository that has been recorded by git as one entry in its log",
  "Commit access", "Ability to make changes to a repository directly, without making a pull request",
  "Contributor", "User that has made at least one commit to a specified repository",
  "Follow", "Add activity by another user to a user's timeline",
  "git", "Open source version control software",
  "GitHub", "Commercial platform that freely hosts git repositories and provides extended collaboration and social networking features, such as pull requests, issues and stars",
  "GitLab", "Open source alternative to GitHub",
  "CodeBerg", "Open source alternative to GitHub",
  "Issue", "Feature of GitHub that records and tracks a bug report, feature request or other suggestion in a repository",
  "Maintainer", "Individual that has overall control of a repository, generally assumed to be its primary contributor. Repositories can have multiple users with commit access in addition to the maintainer.",
  "Merge", "Accept a pull request and incorporate its changes into a repository.",
  "Organization", "Entity representing a group of users, which can also own repositories",
  "Pull request", "Mechanism by which users that don't have commit access to a repository can contribute to it. The repository's maintainer or another user with commit access must decide whether to merge (accept) the changes, or decline them.",
  "Repository", "Individual project that uses git for version control. Can include a mix of different types of files.",
  "Star", "GitHub's version of a 'like', applied by users to a repository.",
  "Timeline", "Chronological feed of GitHub activity from repositories a user has starred and other users they follow. Also includes repositories that a user is not following if they are 'trending' or determined relevant by GitHub's algorithm.",
  "User", "On GitHub, an individual with an account that can own repositories",
  "Version control", "System for tracking changes (additions, alterations, or deletions) in a set of files, typically but not exclusively computer code"
) |>
  gt() |>
  cols_label_with(fn = tools::toTitleCase) |>
  fmt_markdown(definition)
```

<!-- TODO:
* Introduce git, GitHub, etc. and describe briefly how they work for non-users
-->

### 2023-10-12 analysis/_05-collaboration.qmd
Add tbl-glossary

As well as hosting source code, GitHub and other software forges include systems for facilitating collaboration on code and other projects.
The basic collaborative workflow is inherited from git, which allows multiple users to *commit* (see @tbl-glossary for definitions of this and other git terminology used in this section) code to the repository.
A user with *commit access* to a repository can change any of its contents at will, so this is usually reserved for the project maintainer and known, trusted collaborators.
GitHub extends this model with its *pull request* feature, by which any user can offer to contribute code to a repository to which they don't have commit access.
The maintainer may then choose to *merge* (accept) or *decline* the pull request, facilitating contributions from a wider network of collaborators without the need for permission to be sought in advance.

@@

GitHub also facilitates collaboration on broader project management tasks, primarily through its *issues* feature.^[Apart from issues, GitHub has a very wide range of project management and social media-like features, including wikis, discussion forums and 'kanban' boards. We have not analysed the use of these features here.]
Unless a repository's maintainer specifically configures it otherwise, any user can create an issue attached to another user's repository, or comment on an existing issue.
Issues are typically used to log and track bug reports, feature requests, and other comments and suggestions from the project's user base.
GitHub's pull request feature is also implemented via this system – a pull request is a special type of issue.

@@

Another way GitHub users can engage with repositories and other users with social media-like features such as *starring* a repository, adding it to a public or private list, or *following* a user.
These actions populate a timeline of through which users can see recent activity and discover new projects related to those they have interacted with in the past.^[This feature of GitHub's timeline was one of the primary ways we compiled open-archaeo.]
While not constituting a direct form of collaboration, these features can facilitate the formation and maintenance of collaborative networks, in the same way that other social media platforms serve other professional networks.

>>>>

As well as hosting source code, GitHub and other software forges include systems for facilitating collaboration on code and other projects.
The basic collaborative workflow is inherited from git, which allows multiple users to commit (see @tbl-glossary for definitions of this and other git terminology used in this section) code to the repository.
A user with commit access to a repository can change any of its contents at will, so this is usually reserved for the project maintainer and known, trusted collaborators.
GitHub extends this model with its pull request feature, by which any user can offer to contribute code to a repository to which they don't have commit access.
The maintainer may then choose to merge (accept) or decline the pull request, facilitating contributions from a wider network of collaborators without the need for permission to be sought in advance.

@@

GitHub also facilitates collaboration on broader project management tasks, primarily through its issues feature.^[Apart from issues, GitHub has a very wide range of project management and social media-like features, including wikis, discussion forums and 'kanban' boards. We have not analysed the use of these features here.]
Unless a repository's maintainer specifically configures it otherwise, any user can create an issue attached to another user's repository, or comment on an existing issue.
Issues are typically used to log and track bug reports, feature requests, and other comments and suggestions from the project's user base.
GitHub's pull request feature is also implemented via this system – a pull request is a special type of issue.

@@

Another way GitHub users can engage with repositories and other users with social media-like features such as starring a repository, adding it to a public or private list, or following a user.
These actions populate a timeline of through which users can see recent activity and discover new projects related to those they have interacted with in the past.^[This feature of GitHub's timeline was one of the primary ways we compiled open-archaeo.]
While not constituting a direct form of collaboration, these features can facilitate the formation and maintenance of collaborative networks, in the same way that other social media platforms serve other professional networks.
